---
title: "Lab 9"
output: html_document
---

## Neural networks (seeds data)

seeds data set from UCI repository1

Explore the data.
Measurements of geometrical properties of kernels belonging to three different varieties of wheat. A soft X-ray technique and GRAINS package were used to construct all seven, real-valued attributes.

```{r}
library(tidyverse)
seeds <- read.table(
  "https://archive.ics.uci.edu/ml/machine-learning-databases/00236/seeds_dataset.txt"
  )
colnames(seeds) <- c("area", 
                     "perimeter", 
                     "compactness", 
                     "length_of_kernel", 
                     "width_of_kernel",
                     "asy_coeff", 
                     "length_of_kernel_groove", 
                     "Class")
summary(seeds)
cor(dplyr::select(seeds, -Class))
```


```{r}
dim(seeds)
```

```{r}
x <- seeds %>%
  dplyr::select(-Class) %>%
  scale()
```

```{r}
set.seed(1)

seeds_train_index <- seeds %>%
  mutate(ind = 1:nrow(seeds)) %>%
  group_by(Class) %>%
  mutate(n = n()) %>%
  sample_frac(size = .75, weight = n) %>%
  ungroup() %>%
  pull(ind)
```


```{r}
library(nnet)
class_labels <- pull(seeds, Class) %>% 
  class.ind() 
```


```{r}
seeds_train <- x[seeds_train_index, ]
train_class <- class_labels[seeds_train_index,]
seeds_test <- x[-seeds_train_index, ] 
test_class <- class_labels[-seeds_train_index,]
```

```{r}
nn_seeds <- nnet(
  x = seeds_train, 
  y = train_class, 
  size = 4, 
  decay = 0, 
  softmax = TRUE,
  maxit=500
  )
```


```{r}
nn_pred <- predict(nn_seeds, seeds_test, 
                   type="class")

tab_seeds <- table(slice(
  seeds, 
  -seeds_train_index) %>% pull(Class), 
  nn_pred)

1-sum(diag(tab_seeds))/sum(tab_seeds)
```


##Neural networks (Boston data (quantitative response))

Let’s consider housing price data, Boston in MASS package.
Response is quantitative.

```{r}
library(nnet)
library(MASS)
```

```{r}
train_Boston <- sample(
  1:nrow(Boston), 
  nrow(Boston)/2
  )

x <- scale(Boston)
```


```{r}
Boston_train <- x[train_Boston, ]
train_medv <- x[train_Boston, "medv"]
Boston_test <- x[-train_Boston, ] 
test_medv <- x[-train_Boston, "medv"]
```

```{r}
nn_Boston <- nnet(
  Boston_train, 
  train_medv,  
  size=10, 
  decay=1, 
  softmax=FALSE, 
  maxit=1000,
  linout=TRUE
  )
```

```{r}
nn_pred <- predict(
  nn_Boston, 
  Boston_test,
  type="raw"
  )
```


```{r}
plot(test_medv, nn_pred)

mean((test_medv - nn_pred)^2)
```


##CV for NN - Iris data

80%/20% training/test set.

```{r}
library(e1071)
library(cluster)
set.seed(1)

data("iris")

Species <- pull(iris, Species)

xy <- dplyr::select(iris, -Species) %>%
  scale() %>% 
  data.frame() %>% 
  mutate(Species = Species) # scale predictors

iris_train_index <- iris %>%
  mutate(ind = 1:nrow(iris)) %>%
  group_by(Species) %>%
  mutate(n = n()) %>%
  sample_frac(size = .8, weight = n) %>%
  ungroup() %>%
  pull(ind)

iris_train <- slice(xy, iris_train_index)
iris_test <- slice(xy, -iris_train_index)
class_labels <- pull(xy, Species) %>% 
  class.ind() 

iris_nnet1 <- tune.nnet(
  Species~., 
  data = iris_train, 
  size = 1:30, 
  tunecontrol = tune.control(sampling = "cross",cross=5)
  )

head(summary(iris_nnet1))

plot(iris_nnet1)
```










